# Sampling the Imaginary


```{r}
library(tidyverse)
library(rethinking)
```


```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

3.2. Sampling to summarize section of the book
3E1. How much posterior probability lies below p = 0.2?

```{r}
mean( samples < 0.2 )
```

3E2. How much posterior probability lies above p = 0.8?
```{r}
mean( samples > 0.8 )
```


3E3. How much posterior probability lies between p = 0.2 and p = 0.8?
```{r}
mean( samples > 0.2 & samples < 0.8 )
```

all three should be 1

```{r}
sum(mean( samples < 0.2 ), mean( samples > 0.8 ), mean( samples > 0.2 & samples < 0.8 ))
```



3E4. 20% of the posterior probability lies below which value of p?


```{r}
quantile(samples, 0.2)
```

3E5. 20% of the posterior probability lies above which value of p?

```{r}
quantile(samples, 0.8)
```


3E6. Which values of p contain the narrowest interval equal to 66% of the posterior probability?

```{r}
HPDI( samples , prob=0.66 )

tidybayes::mode_hdi(samples, .width = .66)

tidybayes::hdi(samples, .width = .66)
```


3E7. Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?


```{r}
PI(samples, prob = 0.66)
```

